---
title: "Assignment_2"
author: "Bridget Fuller"
date: "06-07-2026"
output:
  html_document:
    df_print: paged
---

```{r}
library(caret)
library(ISLR)
head("UniversalBank")
summary(UniversalBank)
```

creating dummy variables for Education field

```{r}
library(dplyr)
bank$Education <- as.factor(bank$Education)
dummies <- model.matrix(~ Education - 1, data = bank)
head(dummies)
dummy_model <- dummyVars(~ Education, data = bank)
education_dummies <- predict(dummy_model, newdata = bank)
head(education_dummies)
bank <- cbind(bank, education_dummies)
head(bank)
```

```{r}
library(dplyr)

names(bank) <- make.names(names(bank))

m_universalbank<- select(bank,Age,Experience,Income,Family,CCAvg,Education.1,Education.2,Education.3,Mortgage,Securities.Account,CD.Account,Online,CreditCard,Personal.Loan)
```

training data sets

```{r}
set.seed(123)
trainIndex <- createDataPartition(bank$Personal.Loan, p = 0.6, list = FALSE)
train <- bank[trainIndex, ]
test  <- bank[-trainIndex, ]
train.norm.df[,3] <- as.factor(train.norm.df[,3])
valid.norm.df[,3] <- as.factor(valid.norm.df[,3])
valid.norm.df[,3] <- factor(valid.norm.df[,3], levels = levels(train.norm.df[,3]))
```

```{r}
set.seed(15)
Test_Index = createDataPartition(m_universalbank$Personal.Loan,p=0.4, list=FALSE) 
Test_Data = m_universalbank[Test_Index,]
TraVal_Data = m_universalbank[-Test_Index,]
Train_Index = createDataPartition(TraVal_Data$Personal.Loan,p=0.6, list=FALSE)
Train_Data = TraVal_Data[Train_Index,]
Validation_Data = TraVal_Data[-Train_Index,]

summary(Train_Data)
summary(Validation_Data)
summary(Test_Data)
```

```{r}
train.norm.df <- Train_Data
valid.norm.df <- Validation_Data
traval.norm.df <- TraVal_Data
```

```{r}
train.norm.df[, 1:2] <- predict(norm.values, Train_Data[, 1:2]) 
valid.norm.df[, 1:2] <- predict(norm.values, Validation_Data[, 1:2])
traval.norm.df[, 1:2] <- predict(norm.values, traval.norm.df[, 1:2])
test.norm.df <- predict(norm.values, Test_Data[, 1:2])

summary(train.norm.df)
var(train.norm.df[, 1:2])
summary(valid.norm.df)
var(valid.norm.df[, 1:2])
```
modeling k-NN after normalization
```{r}
library(FNN)
nn <- knn(train = train.norm.df[, 1:2], test = test.norm.df, 
          cl = train.norm.df[, 3], k = 1, prob=TRUE) 
row.names(Train_Data)[attr(nn, "nn.index")]
```
tuning K
```{r}
set.seed(123)

model <- train(
  Personal.Loan ~ Age  + Income,
  data = predictors_normalized,
  method = "knn",
  tuneLength = 10
)

model
plot(model)

predictors_normalized$Personal.Loan <- as.factor(predictors_normalized$Personal.Loan)
```
pred <- predict(model, predictors_normalized)
```

final value K=23 

Confusion Matrix for Validation set

```{r}
library("gmodels")
CrossTable(x=Test_labels,y=Predicted_Test_labels, prop.chisq = FALSE)
```

```{r}
Predicted_Test_labels <-knn(Train_Predictors, 
                           Test_Predictors, 
                           cl=Train_labels, k=100, prob=TRUE )

class_prob<-attr(Predicted_Test_labels, 'prob')
head(class_prob)
```

```{r}
bank$Education <- as.factor(bank$Education)
dummies <- model.matrix(~ Education - 1, data = bank)

model_df <- cbind(
  bank[, c("Age","Experience","Income","Family","CCAvg",
           "Mortgage","Securities.Account","CD.Account",
           "Online","CreditCard","Personal.Loan")],
  dummies
)
norm_model <- preProcess(model_df[, -which(names(model_df) == "Personal.Loan")],
                         method = "range")

model_norm <- predict(norm_model, model_df)
model_norm$Personal.Loan <- as.factor(model_df$Personal.Loan)

set.seed(123)
idx <- createDataPartition(model_norm$Personal.Loan, p = 0.8, list = FALSE)

Train <- model_norm[idx, ]
Test  <- model_norm[-idx, ]
```


```{r}
library(class)

Train_X <- Train[, names(new_customer_norm)]
Train_y <- Train$Personal.Loan

prediction <- knn(
  train = Train_X,
  test = new_customer_norm,
  cl = Train_y,
  k = 1
)

prediction
```

question 1 customer prediction 0. will not except the loan

```{r}
head(bank)
bank$Education <- as.factor(bank$Education)
edu_dummies <- model.matrix(~ Education - 1, data = bank)

model_df <- cbind(
  bank[, c("Age","Experience","Income","Family","CCAvg",
           "Mortgage","Securities.Account","CD.Account",
           "Online","CreditCard","Personal.Loan")],
  edu_dummies
)

norm_model <- preProcess(model_df[, -which(names(model_df) == "Personal.Loan")],
                         method = "range")

model_norm <- predict(norm_model, model_df)
model_norm$Personal.Loan <- as.factor(model_df$Personal.Loan)

new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1,
  Education1 = 0,
  Education2 = 1,
  Education3 = 0
)
new_customer_norm <- predict(norm_model, new_customer)

library(class)

Train_X <- Train[, names(new_customer_norm)]
Train_y <- Train$Personal.Loan

prediction <- knn(
  train = Train_X,
  test = new_customer_norm,
  cl = Train_y,
  k = 23
)

prediction
```

prediction for new customer is not to except using best K=23


# 50/30/20 split
```{r}

set.seed(123)
idx_train <- createDataPartition(data$Personal.Loan, p = 0.5, list = FALSE)
Train <- data[idx_train, ]
Temp  <- data[-idx_train, ]

set.seed(123)
idx_valid <- createDataPartition(Temp$Personal.Loan, p = 0.6, list = FALSE)
Valid <- Temp[idx_valid, ]
Test  <- Temp[-idx_valid, ]

# Prepare predictors and labels
Train_X <- Train[, setdiff(names(Train), "Personal.Loan")]
Valid_X <- Valid[, setdiff(names(Valid), "Personal.Loan")]
Test_X  <- Test[,  setdiff(names(Test),  "Personal.Loan")]

Train_y <- Train$Personal.Loan
Valid_y <- Valid$Personal.Loan
Test_y  <- Test$Personal.Loan

# KNN with k = 23
k_val <- 23
pred_train <- knn(Train_X, Train_X, cl = Train_y, k = k_val)
pred_valid <- knn(Train_X, Valid_X, cl = Train_y, k = k_val)
pred_test  <- knn(Train_X, Test_X,  cl = Train_y, k = k_val)

# Confusion matrices
cm_train <- confusionMatrix(pred_train, Train_y)
cm_valid <- confusionMatrix(pred_valid, Valid_y)
cm_test  <- confusionMatrix(pred_test,  Test_y)

cm_train
cm_valid
cm_test






