---
title: "Assignment_4"
author: "Bridget Fuller"
date: "2026-06-14"
output: pdf_document
---

#loading in the data set Pharmaceuticals.cvs and needed libraries install if needed
```{r}
summary(Pharmaceuticals)
library(tidyverse)
library(factoextra)
library(ISLR)
set.seed(123)

df<-Pharmaceuticals[,c(3,11)]
summary(df)
```
# I am using market cap and net profit margin, the data must be scaled to make sure the units match

```{r}
df<-scale(df)
distance<-get_dist(df)
fviz_dist(distance)
```
# this will graph the data
# using K=3 for the number of clusters and visualizing the output
```{r}
k3<-kmeans(df,centers=3,nstart=25)
k3$centers #number of centers
k3$size  #number of companies in cluster
k3$cluster[21] # where is the 21st company
fviz_cluster(k3,data=df)
```
# it is now easier to see the company with the largest market cap and net profit margin. You can see that cluster 1 has the largest values and cluster 3 has the smallest values but the largest population. The highest value companies have recommendations of "hold" or "moderate Buy" no other cluster has this mix.
#Cluster 1  name Large Companies, Cluster# 2 Mid Size Companies, Cluster # 3 Growing Companies 


#other distance checks

```{r}
library(flexclust)
set.seed(123)
k3=kcca(df,k=3,kccaFamily("kmedians"))
k3
```

#now predict the functions
```{r}
cluster_index<-predict(k3)
dist(k3@centers)
image(k3)
points(df,col=cluster_index,pch=19,cex=0.3)
```
#finding K with elbow and silhouette charts
```{r}
library(tidyverse)
library(factoextra)
library(ISLR)
set.seed(123)
df<-scale(df)
fviz_nbclust(df,kmeans,method="wss")
```
# the elbow chart shows K should equal 3
#silhouette chart
```{r}
fviz_nbclust(df,kmeans,method="silhouette")
```
#  this also shows that k=3









